---
title: "Basic Model Run"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic_Model_Run}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dpi = 250,
  fig.width = 7
)
```


## 1. Simplest run

We can get started with the simplest model run with almost entirely default model parameters. The only parameter that malariasimple requires you to define in the initial EIR (init_EIR)

```{r, message = FALSE, results = "hide", fig.align = 'center', out.width='100%', fig.asp=0.55}
library(malariasimple)
#Generate model
gen <- gen_model()

#Define default parameters
params <- get_parameters() |>
  set_equilibrium(init_EIR = 10)

#Run model
out <- run_simulation(gen,params) |> as.data.frame()

#Plot PfPR[2-10]
plot(x = out$day, y = out$n_detect_730_3650/out$n_730_3650,
     type = "l", col = "red", lwd = 2,
     xlab = "Days", ylab = expression(paste(italic(Pf),"PR"[2-10])),
     ylim = c(0,1))
```

## 2. Setting up basic parameters

Simulating a a model in *malariasimple* is a two-step process: first define parameters, then run the model. You can define all core parameters using the helper function `get_parameters()`. For a detailed description of the parameters and their default values, refer to the documentation by running `?get_parameters`.

The other key function in preparing parameters is `set_equilbrium()`. This calculates equilibrium parameters at the specified initial EIR. This function also provides some 'finishing touches' to the parameter list in order to satisfy the model. It is therefore essential that `set_equilbrium()` is called at the end of the parameter preparing process.

```{r, eval = FALSE}
custom_params <- get_parameters(n_days = 200, #Length of simulation
                                ft = 0.3, #Proportion treated
                                eta = 1/(30*365)) |> #Human population birth/death rate
  set_equilibrium(init_EIR = 10)
custom_out <- run_simulation(gen,custom_params)
```

#### Age Group Rendering

From `get_parameters()`, we can define which age groups we want the model to output regarding clinical incidence and prevalence. It is essential that these align with the age compartments defined in age_vector (for default values, see ?get_parameters). All ages must be provided in *days*. To align with `malariasimulation`, prevalence is inferred from the proportion of detected infections (n_detect\_) and the population proportion within that age group (n\_)

In the following examples we will request prevalence and clinical incidence for two age groups:

- Ages 2-10
- Everyone

```{r, message = FALSE, results = "hide", fig.align = 'center', out.width='100%', fig.asp=0.55}
#Define age categories
age_vector <- c(0,1,2,5,10,20,40,60,100)*365
min_age_render <- c(2,0)*365
max_age_render <- c(10,Inf)*365

## Set up model parameters
params_ages <- get_parameters(n_days = 100,
                         age_vector = age_vector,
                         clin_inc_rendering_min_ages = min_age_render,
                         clin_inc_rendering_max_ages = max_age_render,
                         prevalence_rendering_min_ages = min_age_render,
                         prevalence_rendering_max_ages = max_age_render) |>
  set_equilibrium(init_EIR = 10)

## Run model
out_ages <- run_simulation(gen,params_ages) |> as.data.frame()

## Plot
par(mfrow = c(1,2))
plot(x = out_ages$day, y = out_ages$n_clin_inc_0_Inf / out_ages$n_0_Inf,
     type = "l", col = "#F8766D", ylim = c(0,0.0025),lwd = 2,
     xlab = "Days", ylab = "Clinical Incidence")
lines(out_ages$day, out_ages$n_clin_inc_730_3650 / out_ages$n_730_3650,
      col = "#00BFC4",lwd=2)
legend("bottomright", legend = c("All ages", "Ages 2-10"),
       col = c("#F8766D", "#00BFC4"), lty = 1,
       bty = "n", ncol = 1, cex = 0.8, lwd=2)

plot(out_ages$day, out_ages$n_detect_0_Inf / out_ages$n_0_Inf,
     type = "l", col = "#F8766D", ylim = c(0,0.4), lwd=2,
     xlab = "Days", ylab = "Prevalence")
lines(out_ages$day, out_ages$n_detect_730_3650 / out_ages$n_730_3650,
      col = "#00BFC4",lwd=2)
legend("bottomright", legend = c("All ages", "Ages 2-10"),
       col = c("#F8766D", "#00BFC4"), lty = 1,
       bty = "n", ncol = 1, cex = 0.8, lwd=2)
```

## 3. Introducing seasonality

Seasonality can be included in the model using the `set_seasonality()` function. The main argument is the coefficients of a Fourier function (FF) describing the relative carrying capacity of the vector population over time. In this example, we will use default parameters for Thies, Senegal using the [ICDMM package](https://mrc-ide.github.io/deterministic-malaria-model/index). FF coefficients can also be input manually.

```{r, message = FALSE, results = "hide", fig.align = 'center', out.width='100%', fig.asp=0.55}
##------------------ Define seasonal parameters --------------------
admin2 <- "Thies" #Define admin2 region

#Import dataframe of FF coefficients for all regions within the ICDMM database
ff_df <- ICDMM::load_file("admin_units_seasonal.rds")

#Find row defining Thies
admin_row_index <- ICDMM::admin_match(admin_unit = admin2, admin_units_seasonal = ff_df)
admin_row <- ff_df[admin_row_index,]

#Put into format required by malariasimple
ff_coeffs <- list(g0 = admin_row["a0"],
                         g = admin_row[c("a1","a2","a3")],
                         h = admin_row[c("b1","b2","b3")],
                         theta_c = admin_row["theta_c"])

#Floor argument defines minimal value of relative carrying capacity
seasonal_params <- get_parameters(n_days = 365,
                                age_vector = c(0, 1, 3, 5, 10, 20, 40, 60)*365) |>
  set_seasonality(ff_coeffs, floor = 0.002) |>
  set_equilibrium(init_EIR = 10)

##------------------- Run simulation and plot ---------------------
seasonal_output <- run_simulation(gen,seasonal_params) |> as.data.frame()

par(mfrow = c(1,1))
plot(seasonal_output$day, seasonal_output$mv, lwd = 2,
     type = "l", xlab = "Days", "ylab" = "Mosquito Population")
```

We can also view the relative carrying capacity (known in the model as 'theta2') directly using the `get_seasonal_forcing()` function

```{r, message = FALSE, results = "hide", fig.align = 'center', out.width='100%', fig.asp=0.55}
theta2 <- get_seasonal_forcing(ff_coeffs, n_days = 365)
plot(theta2, lwd = 2, type = "l",
     xlab = "Days", ylab = "theta2")
```


